---
title: "magclass for REMINDers"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
date: "2026-07-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dev = "svg", # svg, png,...
  fig.ext = ".svg",
  # fig.asp = .8 # default aspect ratio
  fig.width = 16,
  fig.height = 9,
  dpi = 100
)
```

```{r message=FALSE}
# load the library
library(magclass)
library(madrat)

# loading other helpers used in this tutorial
library(tidyr)
library(dplyr)
library(piamutils)

# load example data
pop <- maxample("pop")
animal <- maxample("animal")
```

# Part 1: Basics

## declaring a magpie object

```{r}
# creating a magpie object with 2 regions, 2 years and 2 different values
m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000),
  names = c("iron", "steel"),
  sets = c("region", "year", "variable"),
  fill = 0
)

# take a look at the data
print(m1)

# creating a magpie object with more than one subdimension
m2 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = seq(1995, 2020, 5),
  names = c("IEA.iron", "IEA.steel", "WoldSteelDatabase.steel"),
  sets = c("region", "year", "model.product"),
  fill = NA
)

# take a look at the data
print(m2)
getSets(m2)

# note that the first and second dimension can also be multidimensional, e.g. for coordinates in dimension one and subannual data in dimension two, but this is not a frequent use case in the REMIND world
```

## creating from a data frame

```{r}
df <- data.frame(
  region = c("USA", "BRA", "FRA"),
  years = c(2000, 2005, 2010),
  product = c("iron", "steel", "cement")
) %>%
  complete(region, years, product) %>%
  mutate("value" = sample.int(n = 1000, size = 27))

print(head(df))

# magpie will try guessing the right attribution. To be safe,
# make the spatial, temporal and data column explicit. All other
# columns will form the third dimension with its subdimensions.
m <- as.magpie(df, spatial = 1, temporal = 2, datacol = 4)
print(m)

# explicitly assigning dimensions is recommended, because otherwise magclass will
# guess dimensions. assume, you have a model column with the model name 'IEA'. as
# it is is a three letter, uppercase model, magclass guesses it is a spatial column

df$model <- "IEA"
as.magpie(df)

# 'as.magpie' seems to struggle with the conversion when the value column is not the
# last column of the data frame. even explicit assignment does not work here
as.magpie(df, spatial = 1, temporal = 2, datacol = 4)


# in order to avoid wrong assignments during the conversion to magpie objects,
# always make sure the column order in your data frame is: regional column, temporal column,
# other columns and the value column at last.
as.magpie(df[, c("region", "years", "model", "product", "value")], spatial = 1, temporal = 2, datacol = 5)
```

## reading in from different file formats

```{r eval=FALSE}
# read from gdx

# if you want to try this yourself, download a release from  https://rse.pik-potsdam.de/data/example/remind2_test-convGDX2MIF_SSP2-EU21-PkBudg650-release.gdx and edit gdxPath to reference the downloaded file
gdxPath <- "./remind2_test-convGDX2MIF_SSP2-EU21-PkBudg650-release.gdx"

# often, setting restoreZeros to FALSE is advisable, to limit the size of the magpie object
m <- gdx2::readGDX(gdxPath, c("vm_prodFe"), field = "l", restoreZeros = FALSE, format = "first_found")

# read from cs4r
cs4rPath <- "./f_fedemandInd.cs4r"
x <- read.magpie(cs4rPath)
```

## writing a magpie object to a file
```{r eval=FALSE}
# write to the cs4r file format
magclass::write.magpie(x = pop, file_name = "population_example.cs4r")

# write to a mif report
# mif format has columns "Model", Scenario", "Region", "Variable", "Unit", so let's
# add all missing information to the magclass object before writing to mif format

popOut <- addDim(pop, dim = 3.2, dimName = "variable", item = "Population (million)")
magclass::write.report(x = popOut, file = "population_example.mif", scenario = "historical", model = "WDI")
```

## exploring a magpie object

```{r}
# show the structure of the object
str(pop)

# show the first elements
head(pop)

# show the last elements
tail(pop)

# which elements are there?
getItems(pop)

# which spatial elements are there?
getItems(pop, dim = 1)

# select dimension via its dimension code.
getItems(pop, dim = 3)

# select dimension via its dimension name.
getItems(pop, dim = "scenario")

# what are the sets (dimensions names) of the data?
getSets(pop)

# what are the years of the data?
getYears(pop)

# print the years as integers
getYears(pop, as.integer = TRUE)

# create a simple plot
mplot(pop)

# investigate a higher dimensional magpie object
animal <- maxample("animal")

getItems(animal)

# list dimension items per subdimension separately
getItems(animal, split = TRUE)

# get all animal species
getNames(animal, dim = 2)

# get all animal colours
getNames(animal, dim = 3)

# the same can be achieved by
getItems(animal, dim = 3.3)

# a magpie object can have several subdimensions in all dimensions, not only the third
getItems(animal, dim = 1)
getItems(animal, dim = 2)
```

## filtering a magpie object

```{r}
# return all data of the A2 scenario for LAM and the years 2005 and 2015
pop["LAM", c(2005, 2015), "A2"]

# return data for regions which have “AS” in its name (pmatch allows for partial matching of the given search string)
pop["AS", , , pmatch = TRUE]

# get all regions except for LAM
pop["LAM", , , invert = TRUE]

# you can to specifically select from dimensions of which you have the dimension name
mselect(pop, scenario = "B1", i = c("FSU", "LAM"))

# or you can use alternatively
pop[list(i = c("FSU", "LAM")), , list(scenario = "B1")]

# filtering for higher dimensionality objects
animal[
  list(x = "5p75", country = c("NLD", "BEL")),
  list(year = "y2000"),
  list(species = "bird", color = "black")
]

# get all birds in Luxembourg
animal["LUX", , "bird"]

# get all red birds
animal[, , "bird.red"]

# this can also be achieved by chaining subsets, i.e.
animal[, , "bird"][, , "red"]

# be careful when using the square bracket filtering method in cases where
# values are not unique to one dimension

# in this example 'ccs' is a possible value for both the technology and scenario dimension

m <- new.magpie(
  cells_and_regions = c("World"),
  years = c(2000, 2005),
  names = c("ccs.ssp2", "ccu.ssp3", "dac.ccs"),
  sets = c("region", "year", "technology", "scenario"),
  fill = sample(6)
)

# this returns two rows, although you might only expect to get one
m[, , "ccs"]

# therefore, this filter is more precise
m[, , list(technology = "ccs")]
```

## manipulating a magpie object

```{r}
pop1 <- pop

# assign 0 to all A2 scenario data
pop1[, , "A2"] <- 0

# overwrite the first region "AFR" with "GHA"
getItems(pop1, dim = 1)[1] <- "GHA"
head(pop1)

# rename the temporal dimension from "t" to "year" and spatial dimension from "i" to "region"
getSets(pop1)[[2]] <- "year"
getSets(pop1)[[1]] <- "region"
getSets(pop1)

# drop all the names in the data dimension that are the same for each element
# in this case, the "type" dimension is dropped, as it is always "animal"
a1 <- collapseDim(animal)
getNames(a1)

# in a2, collapseDim would drop both the "type" and "species" dimension,
# because "type" is always "animal" and "species" always "rabbit"
a2 <- animal[, , "rabbit"]

# this call makes sure that the "type" dimension is not dropped although
# it only contains the value "animal"
a2 <- collapseDim(a2, keepdim = "type")
getNames(a2)

# another way of dropping the type dimension is this
a3 <- animal
getItems(a3, dim = "type") <- NULL

# but careful, dropping a dimension with more than one value this way will introduce duplicates!
getItems(a3, dim = "color") <- NULL

# add a new dimension to existing data
pop1 <- addDim(pop, dim = 3.1, dimName = "model", item = "WDI")
getNames(pop1)

# add a new entry to an existing dimension, set all values to NA
pop1 <- add_columns(pop1, addnm = "C3", dim = 3.2, fill = NA)
getNames(pop1)

# let "C3" values be the difference between A2 and B1
pop1[, , "C3"] <- pop1[, , "A2"] - pop1[, , "B1"]
head(pop1[, , "C3"])

# you can apply mathematic operations to each item of the magclass objects
m3 <- pop["AFR", c(2015, 2025), "B1"]
print(m3)

# for example:

# multiply each item with a factor (scalar)
m3 * 2.5

# squared of each item
m3^2

# this could be used to convert the world population data from millions to billions
pop1 <- pop1 / 1000

# set all values greater than 1 to 1
pop1[pop1 > 1] <- 1
max(pop1)
```

## exercises from the workshop / tutorial

```{r eval=FALSE}
# calculate the share of the global population for each region per year
pop / dimSums(pop, dim = 1)

# take the "animal" example data set from the magclass package and compute how many rabbits there are reported in Luxembourg (LUX) in may of each available year.
collapseDim(dimSums(animal["LUX", "may", "rabbit"], dim = c(3.1, 3.3)))

a1 <- animal

# set all values for red birds in the animal data set to 0 and all values greater than 10 to 10.
a1[, , "bird"][, , "red"] <- 0
a1[a1 > 10] <- 10

# rename all dogs in the animal data set to cats and remove the "type" subdimension.
getNames(a1, dim = 2)[3] <- "cat"
a1 <- collapseDim(a1)
```

# Part 2: Advanced operations

## summing over dimensions

```{r}
# get the global population
dimSums(pop, dim = 1)


# a sample with made-up emissions in various industry sectors
m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement", "chemicals"),
  sets = c("region", "year", "emissions"),
  fill = sample(16)
)

# get emissions over all sectors
dimSums(m, dim = 3)

# get emissions over all sectors and all years
dimSums(m, dim = c(2, 3))

# you can also sum over subdimensions
animals <- maxample("animal")
# sum over colors in animals object
head(dimSums(animals, dim = 3.3))

# be careful when a magpie object contains NAs
m["USA", 2000, "cement"] <- NA
# summing over sectors yields NA for USA in 2000, because cement has no value
# for that year
dimSums(m, dim = 3)

# this behaviour makes sense, as we cannot assume that data gaps can be interpreted
# as 0 values. if you still want to, you can set na.rm = TRUE
dimSums(m, dim = 3, na.rm = TRUE)
```

## looping with 'magpply'

```{r}
m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("cement"),
  sets = c("region", "year", "emissions"),
  fill = c(72, NA, NA, NA)
)

print(m)

# efficient looping over magpie can be done using 'magpply'
# this call loops over the spatial dimension and identifies rows with all NAs
magpply(m, function(y) all(is.na(y)), MARGIN = 1)

# this could be useful when dropping such countries
m[!magpply(m, function(y) all(is.na(y)), MARGIN = 1), , ]

# same could be done for the temporal dimension
magpply(m, function(y) all(is.na(y)), MARGIN = 2)
```

## multiplication of two magpie objects
```{r}
m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "product"),
  fill = sample(12)
)

print(m)

# simple case: all dimension values in "factor" either perfectly match pop1 (year), or
# they have exactly one value (region, product)
factor <- new.magpie(years = getItems(m, 2), fill = c(1.5, 2, 4))

print(factor)

# if you do this calculation, the factors will be multiplied with the matching years
# the same factors are applied for each region and scenario
m * factor

# behind the scenes, factor is expanded to match m to then do an element-wise multiplication
magpie_expand(factor, m)

# what happens if the "factor" object does not have matching years?
# magpie assumes the years in both objects are of different sets and creates the cross product between all years in both sets. therefore, the object is blown up in size
m * factor[, c(1995, 2005), ]

# if you do calculations with two magpie objects with exact same dimensions, the data
# will be applied element-wise

factor <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  fill = seq(1, 2, length.out = 12)
)

print(factor)

m * factor
```

## merging two magpie objects

```{r}
m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "product"),
  fill = sample(12)
)

m2 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("chemicals"),
  sets = c("region", "year", "product"),
  fill = sample(4)
)

# appending one magpie object to another one is done using 'mbind'

mbind(m1, m2)

# however, binding two objects is not straightforward when the dimensions do not match
# the following will crash, because both temporal and spatial dimension differ

m3 <- new.magpie(
  cells_and_regions = c("DEU", "FRA"),
  years = c(2000, 2005),
  names = c("chemicals"),
  sets = c("region", "year", "product"),
  fill = sample(4)
)

# this will crash
# mbind(m1, m3)

# one way of fixing this is adding missing dimension values before binding, but this can
# be cumbersome for larger differences
m1 <- add_columns(m1, addnm = c("DEU", "FRA"), dim = 1, fill = NA)
m3 <- add_columns(m3, addnm = c("y1995"), dim = 2, fill = NA)
m3 <- add_columns(m3, addnm = c("USA", "BRA"), dim = 1, fill = NA)

mbind(m1, m3)

# in some cases, the best way to deal with this is defining a new magpie object
# that has all the dimension values and copy over values

m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "product"),
  fill = sample(12)
)

m2 <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(2010, 2015),
  names = c("chemicals"),
  sets = c("region", "year", "product"),
  fill = sample(4)
)

m3 <- new.magpie(
  cells_and_regions = union(getItems(m1, dim = 1), getItems(m2, dim = 1)),
  years = union(getItems(m1, dim = 2), getItems(m2, dim = 2)),
  names = union(getNames(m1), getNames(m2)),
  fill = NA
)

m3[getItems(m1, dim = 1), getItems(m1, dim = 2), getNames(m1)] <- m1
m3[getItems(m2, dim = 1), getItems(m2, dim = 2), getNames(m2)] <- m2
m3

# sometimes you just want to fit one magpie object to a reference to
# make appending possible
m1 <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(1995, 2000, 2005),
  names = c("chemicals"),
  sets = c("region", "year", "sector"),
  fill = sample(6)
)

m2 <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(2000, 2005, 2010),
  names = c("steel", "cement"),
  sets = c("region", "year", "sector"),
  fill = sample(12)
)

# this matches the years of m1 to the years of m2, i.e. 1995 is dropped and
# 2005 is added with NAs. afterwards, the two magpie object can be merged into one
m1 <- matchDim(x = m1, ref = m2, dim = 2, fill = NA)
mbind(m1, m2)

# matchDim can also be used to make one magpie object match another one
# so you can do element-wise mutliplication

m1 <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(1995, 2000, 2005, 2010),
  names = c("steel"),
  sets = c("region", "year", "sector"),
  fill = sample(6)
)

factor <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(2000, 2005, 2010, 2015),
  names = c("steel", "chemicals"),
  sets = c("region", "year", "sector"),
  fill = sample(12)
)

# this makes m1 match factor by removing 1995 from m1, adding 2015 with all 0s and
# adding the sector 'chemicals' with all 0s
# as 'dim' is not specified, all dimensions are matched
# afterwards, an element-wise multiplication becomes possible
# added with all 0s
m2 <- matchDim(m1, factor, fill = 0)
m2 * factor
```

## time interpolation

```{r}
m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2005),
  names = NULL,
  sets = c("region", "year", "emissions"),
  fill = c(72, 81, 99, 102)
)

print(m)

# you can fill missing time steps by linear interpolation, e.g. add a 2000 timestep
time_interpolate(m, interpolated_year = seq(1995, 2005, 5))

# you can also extrapolate timesteps before or after the time steps
time_interpolate(m, interpolated_year = seq(1990, 2010, 5), extrapolation_type = "linear")

# you can specify that extrapolated time steps just repeat the earliest / latest value
time_interpolate(m, interpolated_year = seq(1990, 2010, 5), extrapolation_type = "constant")

# the current extrapolation method has limitations, though ..

m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("cement", "steel"),
  sets = c("region", "year", "emissions"),
  fill = c(72, 81, NA, 99, 102, NA, 80, NA, 90, 12, NA, 33)
)

# take this magpie object
print(m)

# interpolation won't fill the NA gaps, because interpolated steps are inferred by missing
# values in the temporal dimension (i.e y2000 has to be missing to create an interpolated step)
# however, this object has NAs for some of the existing time steps for some regions and sectors
time_interpolate(m, interpolated_year = seq(1995, 2005, 5))

# for this to work, you'd have to do row-wise interpolation, dropping NAs from the rows
# this is admittedly a bit cumbersome
for (reg in getItems(m, 1)) {
  for (sec in getItems(m, 3)) {
    dropYears <- magpply(m[reg, , sec], function(y) all(is.na(y)), MARGIN = 2)
    m[reg, , sec] <- time_interpolate(m[reg, !dropYears, sec], interpolated_year = seq(1995, 2005, 5))
  }
}

print(m)
```


# Part 3: Useful tool functions

## managing ISO countries

```{r eval=FALSE} 
# retrieve a list of all 249 ISO countries with their full names and 3-letter codes
madrat::getISOlist()

# add all the missing ISO countries to a magpie object
m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995),
  names = c("cement"),
  sets = c("region", "year", "emissions"),
  fill = sample(2)
)

# new values are set to NA, warnings are not printed
madrat::toolCountryFill(m, fill = NA, verbosity = 2)

# regions that are not ISO countries are dropped, throwing a warning
# 'Data for following unknown country codes removed: EUR'

m <- new.magpie(
  cells_and_regions = c("USA", "EUR"),
  years = c(1995),
  names = c("cement"),
  sets = c("region", "year", "emissions"),
  fill = sample(2)
)

# you can list regions that you intentionally want to drop to omit that warning
madrat::toolCountryFill(m, fill = NA, verbosity = 2, no_remove_warning = c("EUR"))

# convert the country names in the spatial region to ISO-3 country codes

m <- new.magpie(
  cells_and_regions = c("United States of America", "England", "Germany"),
  years = c(1995),
  names = c("cement"),
  sets = c("region", "year", "emissions"),
  fill = sample(4)
)

# not all names can be mapped automatically and will produce NA
# 'Following country names could not be found in the country list and returned as NA: "England"'
toolCountry2isocode(getItems(m, dim = 1))

# turn off the warning for England
toolCountry2isocode(getItems(m, dim = 1), ignoreCountries = c("England"))

# .. or add manual mapping entries
toolCountry2isocode(getItems(m, dim = 1), mapping = c("England" = "GBR"))

# adjusting the regions most be done explicitly. do not ignore warnings produced in this step
getItems(m, dim = 1) <- toolCountry2isocode(getItems(m, dim = 1), mapping = c("England" = "GBR"))
```

## (dis)aggregating dimensions

```{r}
# for renaming the values of a single dimension, you can use 'toolAggregate'

m <- new.magpie(
  cells_and_regions = c("FRA", "USA"),
  years = c(2000, 2005, 2010),
  names = c("steel.automobile", "cement.B&C", "chemicals.agriculture", "chemicals.other"),
  sets = c("region", "year", "product.sector"),
  fill = sample(24)
)

print(m)

# in this example, we want to rename the sector names
sectormap <- data.frame(
  from = c("automobile", "B&C", "agriculture", "other"),
  to = c("transport", "buildings", "other", "other")
)

# note that 'agriculture'  and ' other' map to other, so values are added up
toolAggregate(x = m, rel = sectormap, dim = 3.2)

# 'toolaggregate' is also used to aggregate and disaggregate regions
# aggregation of regions works as in the example before 
# the following example shows how disaggregation works

m <- new.magpie(
  cells_and_regions = c("CAZ", "USA", "IND"),
  years = c(2000, 2005, 2010),
  names = c("co2"),
  sets = c("region", "year", "emissions"),
  fill = sample(9)
)

print(m)

# normally, such a regionmapping would be retrieved using a call like this
# toolGetMapping("regionmappingH12.csv"), assuming madrat is set up correctly

regionmapping <- data.frame(
  RegionCode = c("CAZ", "CAZ", "CAZ", "USA", "IND"),
  CountryCode = c("CAN", "NZL", "AUS", "USA", "IND")
)

# the disaggregation needs a weight as magpie object according to which
# the values in CAZ should be split into CAN, NZL, AUS. 
# here we use a 50%/20%/30% ratio
# note that weights do not have to be normalized
w <- data.frame(
  to = c("CAN", "NZL", "AUS", "USA", "IND"),
  weight = c(0.5, 0.2, 0.3, 1, 1)
) |> as.magpie()

toolAggregate(x = m, rel = regionmapping, dim = 1, from = "RegionCode", to = "CountryCode", weight = w)

```


## disaggregating historic regions

```{r message=FALSE}
# this is a special case of disaggregation country where a historic regions (e.g. Former Soviet Union or Yugoslavia) has values until a specific year, then ceases to exist and the values are reported for the individual countries that followed

# if you need to disaggregate the values for the historic country to the current countries, 'toolISOhistorical' can help

# the magpie object has data for Yugoslavia until 1989, after that the data
# is reported for the former Yugoslavian countries individually
m <- new.magpie(
  cells_and_regions = c("YUG", "SRB", "MNE", "SVN", "HRV", "MKD", "BIH"),
  years = seq(1985, 1995, 1),
  names = c("steel"),
  sets = c("region", "year", "emissions"),
  fill = NA
)
m["YUG", seq(1985, 1989, 1), ] <- sample(5)
m[c("SRB", "MNE", "SVN", "HRV", "MKD", "BIH"), seq(1990, 1995, 1), ] <- sample(36)

print(m)

histMap <- data.frame(
  fromISO = "YUG",
  toISO = c("SRB", "MNE", "SVN", "HRV", "MKD", "BIH"),
  lastYear = "y1989"
)

# this call disaggregates YUG values until 1989 and removes YUG, using
# the values in 1990 as a weight
madrat::toolISOhistorical(m, mapping = histMap)
```

## comparing magpie objects
```{r}
m1 <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "sector"),
  fill = sample(12)
)

m2 <- m1
m2["FRA", 2000, ] <- m2["FRA", 2000, ] * 0.2

# sometimes you want to compare the differences between two magpie objects, for
# example to make sure they are still identical after you made some changes

piamutils::compareMagpieObject(m1, m2)

# compareMagpieObject accepts either magpie objects or paths to files containing
# magpie objects
```


# Sources
- [magclass concept: Explaining the general concept behing magclass objects in comparison to arrays and data frames](https://pik-piam.r-universe.dev/magclass/doc/magclass-concept.html)
- [magclass main tutorial: Central tutorial explaining the basics how to use magclass](https://pik-piam.r-universe.dev/magclass/doc/magclass.html)
- [magclass 6: Tutorial about some important changes which came with the complete rewrite of the package in version 6.](https://pik-piam.r-universe.dev/magclass/doc/magclass6.html)
- [magclass object expansion: more specific tutorial going into the details of object expansion and set matching.](https://pik-piam.r-universe.dev/magclass/doc/magclass-expansion.html)
- [The magclass tutorial](https://magpiemodel.github.io/tutorials/t13-magclass)
- [magclass session as part of the REMIND-MAgPIE workshop](https://pik-potsdam.zoom-x.de/rec/share/HXDw1zHT0Vi-uQroAd7GTr1CBbpgng0C8uESaDCcnhfU2RCZHb_do4_qGnBlvXfs.ubV6oTJ__6CioLkt): use code `Ya&1Yi*v`
