---
title: "magclass for REMINDers"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
date: "2026-07-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dev = "svg", # svg, png,...
  fig.ext = ".svg",
  # fig.asp = .8 # default aspect ratio
  fig.width = 16,
  fig.height = 9,
  dpi = 100
)
```

```{r message=FALSE}
# load the library
library(magclass)

# loading other helpers used in this tutorial
library(tidyr)
library(dplyr)

# load sample data used
pop <- maxample("pop")
animal <- maxample("animal")
```

# Part 1: Basics

## declaring a magpie object

```{r}
# creating a magpie object with 2 regions, 2 years and 2 different values
m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000),
  names = c("iron", "steel"),
  sets = c("region", "year", "variable"),
  fill = 0
)

# take a look at the data
print(m1)

# creating a magpie object with more than one subdimension
m2 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = seq(1995, 2020, 5),
  names = c("IEA.iron", "IEA.steel", "WoldSteelDatabase.steel"),
  sets = c("region", "year", "model.product"),
  fill = NA
)

# take a look at the data
print(m2)
getSets(m2)

# note that the first and second dimension can also be multidimensional, e.g. for coordinates in dimension one and subannual data in dimension two, but this is not a frequent use case in the REMIND world
```

## creating from a data frame

```{r}
df <- data.frame(
  region = c("USA", "BRA", "FRA"),
  years = c(2000, 2005, 2010),
  product = c("iron", "steel", "cement")
) %>%
  complete(region, years, product) %>%
  mutate("value" = sample.int(n = 1000, size = 27))

print(head(df))

# magpie will try guessing the right attribution. To be safe,
# make the spatial, temporal and data column explicit. All other
# columns will form the third dimension with its subdimensions.
m <- as.magpie(df, spatial = 1, temporal = 2, datacol = 4)
print(m)
```

## reading in a magpie object from different file formats

```{r eval=FALSE}
# read from gdx

# if you want to try this yourself, download a release from  https://rse.pik-potsdam.de/data/example/remind2_test-convGDX2MIF_SSP2-EU21-PkBudg650-release.gdx and edit gdxPath to reference the downloaded file
gdxPath <- "./remind2_test-convGDX2MIF_SSP2-EU21-PkBudg650-release.gdx"

# often, setting restoreZeros to FALSE is advisable, to limit the size of the magpie object
m <- gdx2::readGDX(gdxPath, c("vm_prodFe"), field = "l", restoreZeros = FALSE, format = "first_found")

# read from cs4r
cs4rPath <- "./f_fedemandInd.cs4r"
x <- read.magpie(cs4rPath)
```

## writing a magpie object to a file
```{r eval=FALSE}
# write to the cs4r file format
magclass::write.magpie(x = pop, file_name = "population_example.cs4r")

# write to a mif report
# mif format has columns "Model", Scenario", "Region", "Variable", "Unit", so let's
# add all missing information to the magclass object before writing to mif format

popOut <- addDim(pop, dim = 3.2, dimName = "variable", item = "Population (million)")
magclass::write.report(x = popOut, file = "population_example.mif", scenario = "historical", model = "WDI")
```

## exploring a magpie object

```{r}
# show the structure of the object
str(pop)

# show the first elements
head(pop)

# show the last elements
tail(pop)

# which elements are there?
getItems(pop)

# which spatial elements are there?
getItems(pop, dim = 1)

# select dimension via its dimension code.
getItems(pop, dim = 3)

# select dimension via its dimension name.
getItems(pop, dim = "scenario")

# what are the sets (dimensions names) of the data?
getSets(pop)

# what are the years of the data?
getYears(pop)

# print the years as integers
getYears(pop, as.integer = TRUE)

# create a simple plot
mplot(pop)

# investigate a higher dimensional magpie object
getItems(animal)

# list dimension items per subdimension separately
getItems(animal, split = TRUE)

# get all animal species
getNames(animal, dim = 2)

# get all animal colours
getNames(animal, dim = 3)

# the same can be achieved by
getItems(animal, dim = 3.3)

# a magpie object can have several subdimensions in all dimensions, not only the third
getItems(animal, dim = 1)
getItems(animal, dim = 2)
```

## filtering a magpie object

```{r}
# return all data of the A2 scenario for LAM and the years 2005 and 2015
pop["LAM", c(2005, 2015), "A2"]

# return data for regions which have “AS” in its name (pmatch allows for partial matching of the given search string)
pop["AS", , , pmatch = TRUE]

# get all regions except for LAM
pop["LAM", , , invert = TRUE]

# you can to specifically select from dimensions of which you have the dimension name
mselect(pop, scenario = "B1", i = c("FSU", "LAM"))

# or you can use alternatively
pop[list(i = c("FSU", "LAM")), , list(scenario = "B1")]

# filtering for higher dimensionality objects
animal[
  list(x = "5p75", country = c("NLD", "BEL")),
  list(year = "y2000"),
  list(species = "bird", color = "black")
]

# get all birds in Luxembourg
animal["LUX", , "bird"]

# get all red birds
animal[, , "bird.red"]

# this can also be achieved by chaining subsets, i.e.
animal[, , "bird"][, , "red"]
```

## manipulating a magpie object

```{r}
pop1 <- pop

# assign 0 to all A2 scenario data
pop1[, , "A2"] <- 0

# overwrite the first region "AFR" with "GHA"
getItems(pop1, dim = 1)[1] <- "GHA"
head(pop1)

# rename the temporal dimension from "t" to "year" and spatial dimension from "i" to "region"
getSets(pop1)[[2]] <- "year"
getSets(pop1)[[1]] <- "region"
getSets(pop1)

# drop all the names in the data dimension that are the same for each element
# in this case, the "type" dimension is dropped, as it is always "animal"
a1 <- collapseDim(animal)
getNames(a1)

# in a2, collapseDim would drop both the "type" and "species" dimension,
# because "type" is always "animal" and "species" always "rabbit"
a2 <- animal[, , "rabbit"]

# this call makes sure that the "type" dimension is not dropped although
# it only contains the value "animal"
a2 <- collapseDim(a2, keepdim = "type")
getNames(a2)

# add a new dimension to existing data
pop1 <- addDim(pop, dim = 3.1, dimName = "model", item = "WDI")
getNames(pop1)

# add a new entry to an existing dimension, set all values to NA
pop1 <- add_columns(pop1, addnm = "C3", dim = 3.2, fill = NA)
getNames(pop1)

# let "C3" values be the difference between A2 and B1
pop1[, , "C3"] <- pop1[, , "A2"] - pop1[, , "B1"]
head(pop1[, , "C3"])

# you can apply mathematic operations to each item of the magclass objects
m3 <- pop["AFR", c(2015, 2025), "B1"]
print(m3)

# for example:

# multiply each item with a factor (scalar)
m3 * 2.5

# squared of each item
m3^2

# this could be used to convert the world population data from millions to billions
pop1 <- pop1 / 1000

# set all values greater than 1 to 1
pop1[pop1 > 1] <- 1
max(pop1)
```

# Part 2: Advanced operations

## multiplication of two magpie objects
```{r}
m <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "product"),
  fill = sample(12)
)

print(m)

# simple case: all dimension values in "factor" either perfectly match pop1 (year), or
# they have exactly one value (region, product)
factor <- new.magpie(years = getItems(m, 2), fill = c(1.5, 2, 4))

print(factor)

# if you do this calculation, the factors will be multiplied with the matching years
# the same factors are applied for each region and scenario
m * factor

# behind the scenes, factor is expanded to match m to then do an element-wise multiplication
magpie_expand(factor, m)

# what happens if the "factor" object does not have matching years?
# magpie assumes the years in both objects are of different sets and creates the cross product between all years in both sets. therefore, the object is blown up in size
m * factor[, c(1995, 2005), ]

# if you do calculations with two magpie objects with exact same dimensions, the data
# will be applied element-wise

factor <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  fill = seq(1, 2, length.out = 12)
)

print(factor)

m * factor
```

## merging two magpie objects

```{r}

m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "product"),
  fill = sample(12)
)

m2 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("chemicals"),
  sets = c("region", "year", "product"),
  fill = sample(4)
)

# appending one magpie object to another one is done using 'mbind'

mbind(m1, m2)

# however, binding two objects is not straightforward when the dimensions do not match
# the following will crash, because both temporal and spatial dimension differ

m3 <- new.magpie(
  cells_and_regions = c("DEU", "FRA"),
  years = c(2000, 2005),
  names = c("chemicals"),
  sets = c("region", "year", "product"),
  fill = sample(4)
)

# this will crash
# mbind(m1, m3)

# one way of fixing this is adding missing dimension values before binding, but this can 
# be cumbersome for larger differences
m1 <- add_columns(m1, addnm = c("DEU", "FRA"), dim = 1, fill = NA)
m3 <- add_columns(m3, addnm = c("y1995"), dim = 2, fill = NA)
m3 <- add_columns(m3, addnm = c("USA", "BRA"), dim = 1, fill = NA)

mbind(m1, m3)

# in some cases, the best way to deal with this is defining a new magpie object
# that has all the dimension values and copy over values

m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000, 2005),
  names = c("steel", "cement"),
  sets = c("region", "year", "product"),
  fill = sample(12)
)

m2 <- new.magpie(
  cells_and_regions = c("FRA", "DEU"),
  years = c(2010, 2015),
  names = c("chemicals"),
  sets = c("region", "year", "product"),
  fill = sample(4)
)

m3 <- new.magpie(
  cells_and_regions = union(getItems(m1, dim = 1), getItems(m2, dim = 1)),
  years = union(getItems(m1, dim = 2), getItems(m2, dim = 2)),
  names = union(getNames(m1), getNames(m2)),
  fill = NA
)
  
m3[getItems(m1, dim = 1),getItems(m1, dim = 2), getNames(m1)] <- m1
m3[getItems(m2, dim = 1),getItems(m2, dim = 2), getNames(m2)] <- m2
m3

```

############ Continue here



```{r}
# summing over dimensions

# get the global population
dimSums(pop, dim = 1)

# TODO add example where we sum over the data dimension

# TODO add example where we sum over the temporal dimension

# TODO sum over multiple dimensions at once

# TODO add example where na.rm = TRUE is needed

# TODO time interpolations
# time_interpolate(d[, , 1], 2005:2030)
```


## Exercises from the Workshop / Tutorial

TODO: add more difficult ones from the magclass tutorial

```{r}
# calculate the share of the global population for each region per year

pop / dimSums(pop, dim = 1)

# compute how many rabbits there are reported LUX in may each year

collapseDim(dimSums(animal["LUX", "may", "rabbit"], dim = c(3.1, 3.3)))
```


# Advanced

## Common problems when converting to magclass
- matchdim
- time interpolation
- mbind and its subtleties
- toolAggregate
  aggregate regions, simple renamings of dimension names, ...
- toolIso2Country
- toolISOhistorical
- subleties / problems
  - wrong assignment of region for three letter columns (e.g. model = UBA)
  - data frames with odd sorting fail to be converted into right magclass structure
  - filtering for items that appear in more than one subdimension might lead to unexpected results
- comparing magclass objects
```{r}
```

```{r}
```

```{r}
```


# Sources
- [magclass concept: Explaining the general concept behing magclass objects in comparison to arrays and data frames](https://pik-piam.r-universe.dev/magclass/doc/magclass-concept.html)
- [magclass main tutorial: Central tutorial explaining the basics how to use magclass](https://pik-piam.r-universe.dev/magclass/doc/magclass.html)
- [magclass 6: Tutorial about some important changes which came with the complete rewrite of the package in version 6.](https://pik-piam.r-universe.dev/magclass/doc/magclass6.html)
- [magclass object expansion: more specific tutorial going into the details of object expansion and set matching.](https://pik-piam.r-universe.dev/magclass/doc/magclass-expansion.html)
- [The magclass tutorial](https://magpiemodel.github.io/tutorials/t13-magclass)
- [magclass session as part of the REMIND-MAgPIE workshop](https://pik-potsdam.zoom-x.de/rec/share/HXDw1zHT0Vi-uQroAd7GTr1CBbpgng0C8uESaDCcnhfU2RCZHb_do4_qGnBlvXfs.ubV6oTJ__6CioLkt): use code `Ya&1Yi*v`
