---
title: "magclass for REMINDers"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
date: "2026-07-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dev = "svg", # svg, png,...
  fig.ext = ".svg",
  # fig.asp = .8 # default aspect ratio
  fig.width = 16,
  fig.height = 9,
  dpi = 100
)
```

```{r message=FALSE}
# load the library
library(magclass)

# loading other helpers used in this tutorial
library(tidyr)
library(dplyr)

# load sample data
pm <- maxample("pop")
a <- maxample("animal")
```

# Part 1: Basics

## declaring a magpie object

```{r}
# creating a magpie object with 2 regions, 2 years and 2 different values
m1 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = c(1995, 2000),
  names = c("iron", "steel"),
  sets = c("region", "year", "value"),
  fill = 0
)

# take a look at the data
print(m1)

# creating a magpie object with more than one subdimension
m2 <- new.magpie(
  cells_and_regions = c("USA", "BRA"),
  years = seq(1995, 2020, 5),
  names = c("IEA.iron", "IEA.steel", "WoldSteelDatabase.steel"),
  sets = c("region", "year", "value"),
  fill = NA
)

# take a look at the data
print(m2)
```

## exploring a magpie object

```{r}
# Show the structure of the object
str(pm)

# Show the first elements
head(pm)

# Show the last elements
tail(pm)

# Which elements are there?
getItems(pm)

# Which spatial elements are there?
getItems(pm, dim = 1)

# Select dimension via its dimension code.
getItems(pm, dim = 3)

# Select dimension via its dimension name.
getItems(pm, dim = "scenario")

# What are the sets (dimensions names) of the data?
getSets(pm)

# What are the years of the data?
getYears(pm)

# Print the years as integers
getYears(pm, as.integer = TRUE)

# Create a simple plot
mplot(pm)

# Investigate a higher dimensional magpie object
getItems(a)

# get all types
getNames(a, dim = 2)

# get all the species
getNames(a, dim = 3)

# the same can be achieved by
getItems(a, dim = 3.3)

# a magpie object can have several subdimensions in all dimensions, not only the third
getItems(a, dim = 1)
getItems(a, dim = 2)
```

## filtering a magpie object

```{r}
# Return all A2 related data for LAM and the years 2005 and 2015
pm["LAM", c(2005, 2015), "A2"]

# Return data for regions which have “AS” in its name (pmatch allows for partial matching of the given search string)
pm["AS", , , pmatch = TRUE]

# Get all regions except for LAM
pm["LAM", , , invert = TRUE]

# If you want to specifically select from one dimension from which you have the dimension name:
mselect(pm, scenario = "B1", i = c("FSU", "LAM"))

# Or you can use alternatively:
pm[list(i = c("FSU", "LAM")), , list(scenario = "B1")]

# Filtering for higher dimensionality objects
a[
  list(x = "5p75", country = c("NLD", "BEL")),
  list(year = "y2000"),
  list(species = "bird", color = "black")
]
```

## manipulating a magpie object's structure

```{r}
pm1 <- pm

# overwrite the first region "AFR" with "GHA"
getItems(pm1, dim = 1)[1] <- "GHA"
print(pm1)

# rename the temporal dimension from "t" to "year"
getSets(pm1)[2] <- "year"
print(pm1)

# drop all the names in the data dimension that are the same for each element
# in this case, the "type" dimension is dropped, as it is always "animal"
a1 <- collapseDim(a)
getNames(a1)

# in a2, collapseDim would drop both the "type" and "species" dimension,
# because "type" is always "animal" and "species" always "rabbit"
a2 <- a[, , "rabbit"]

# this call makes sure that the "type" dimension is not dropped although
# it only contains the value "animal"
a2 <- collapseDim(a2, keepdim = "type")
getNames(a2)

# add a new dimension to existing data
pm1 <- add_dimension(pm, dim = 3.1, add = "model", nm = "WEO")
getNames(pm1)

# add a new entry to an existing dimension
pm1 <- add_columns(pm1, addnm = "C3", dim = 3.2, fill = NA)
getNames(pm1)

# let "C3" values be the difference between A2 and B1
pm1[, , "C3"] <- pm1[, , "A2"] - pm1[, , "B1"]
head(pm1[, , "C3"])

# you can apply mathematic operations on each item of the magclass objects
m3 <- pm["AFR", c(2015, 2025), "B1"]
print(m3)

# for example:

# multiply each item with a factor
m3 * 2.5

# squared of each item  
m3 ^ 2 



```

## creating from a data frame

```{r}
df <- data.frame(
  region = c("USA", "BRA", "FRA"),
  years = c(2000, 2005, 2010),
  product = c("iron", "steel", "cement")
) %>%
  complete(region, years, product) %>%
  mutate("value" = sample.int(n = 1000, size = 27))

print(head(df))

# magpie will try guessing the right attribution. To be safe,
# make the spatial, temporal and data column explicit. All other
# columns will form the third dimension with its subdimensions.
m <- as.magpie(df, spatial = 1, temporal = 2, datacol = 4)
print(m)
```

- dimsums

## Advanced

- toolAggregate
  aggregate regions, simple renamings of dimension names, ...
- toolIso2Country
- toolISOhistorical
- subleties
- comparing magclass objects
- reading and writing magclass objects




```{r}
```


```{r}
```

```{r}
```

```{r}
```
